---
output: html_document
---

<link rel="stylesheet" href="https://maxcdn.bootstrapcdn.com/bootstrap/3.3.7/css/bootstrap.min.css">
<link rel="stylesheet" href="./custom.css">

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(magrittr)
library(zoo)
library(knitr)
```

<div class='jumbotron'>
  <h2 class='display-3 text-uppercase'>Project 2 - Data Set 2</h2>
  <h4 class='right text-uppercase'>By Brian Weinfeld</h4>
  <div class='clearfix'></div>
  <h5 class='right text-uppercase'>March 3rd, 2018</h5>
</div>

<div class='page-header text-uppercase'>
  <h3>World Populations</h3>
</div>

<div class='well'>
For my second ata set, I used the set provided by Rose. This data set identifies a series of regions in the United States and, for each day in January 2018, shows the amount of elecriticy generated, the amount consumed, and the difference between these values.
</div>

```{r, warning=FALSE, message=FALSE}
raw.data <- read_csv('C:\\Users\\Brian\\Desktop\\GradClasses\\Spring18\\migrantdata\\us_daily_electric_system_operating_data.csv', 
                     skip=4)
kable(raw.data[, 1:5] %>% head(10))
```

<div class='well'>
In order to tidy this data, I would first need to extract the locations present in every fourth row and create a new column in the original data set in which to store the information. Annoyingly, one of the location names features a comma and thus half the name is placed in the 01/01/2018 column. I developed a generic solution to this issue in case it came up in a bigger data set.
</div>

```{r}
locations <- raw.data[seq(1, length(raw.data), 4), 1:2] %>% 
  replace_na(list(`01/01/2018`='')) %>%
  unite(location, 1, 2, sep='') %>%
  select(1) %>%
  unlist()
kable(locations)
```

<div class='well'>

</div>

















