---
output: html_document
---

<link rel="stylesheet" href="https://maxcdn.bootstrapcdn.com/bootstrap/3.3.7/css/bootstrap.min.css">
<link rel="stylesheet" href="./custom.css">

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(magrittr)
library(zoo)
library(knitr)
```

<div class='jumbotron'>
  <h2 class='display-3 text-uppercase'>Project 2 - Data Set 2</h2>
  <h4 class='right text-uppercase'>By Brian Weinfeld</h4>
  <div class='clearfix'></div>
  <h5 class='right text-uppercase'>March 3rd, 2018</h5>
</div>

<div class='page-header text-uppercase'>
  <h3>Electricity Demand</h3>
</div>

<div class='well'>
For my second data set, I used the set provided by Rose. This data set identifies a series of regions in the United States and, for each day in January 2018, shows the amount of elecriticy generated, the amount consumed, and the difference between these values.
</div>

```{r, warning=FALSE, message=FALSE}
raw.data <- read_csv('C:\\Users\\Brian\\Desktop\\GradClasses\\Spring18\\migrantdata\\us_daily_electric_system_operating_data.csv', 
                     skip=4)
kable(raw.data[, 1:5] %>% head(10))
```

<div class='well'>
In order to tidy this data, I would first need to extract the locations present in every fourth row and create a new column in the original data set in which to store the information. Annoyingly, one of the location names features a comma and thus half the name is placed in the 01/01/2018 column. I developed a generic solution to this issue in case it came up in a bigger data set.
</div>

```{r}
locations <- raw.data[seq(1, length(raw.data), 4), 1:2] %>% 
  replace_na(list(`01/01/2018`='')) %>%
  unite(location, 1, 2, sep='') %>%
  select(1) %>%
  unlist()
kable(locations)
```

<div class='well'>
With all the necessary information, I tidied up the data and made a few additional changes to make this particular data set easier to work with. That included changing the names of some columns and splitting the date into a month, day and year column. 
</div>

```{r}
tidy.data <- raw.data[which(1:length(raw.data) %% 4 != 1), ] %>% 
  mutate(location=locations %>% 
              rep(each=3)
            ) %>%
  gather(date, mega.watt.hours, 2:32) %>%
  spread(megawatthours, mega.watt.hours) %>%
  rename(generated=`Net generation`, net=`Total net actual interchange`, demand=Demand) %>%
  separate(date, c('month', 'day', 'year'))

tidy.data[, 2:7] %<>%
  map(~as.numeric(.))
kable(tidy.data %>% head(10))
```

<div class='page-header text-uppercase'>
  <h3>Analysis</h3>
</div>

<div class='well'>
I plotted the proportion of energy generated compared to demand for each day and location. Any points above 1.0 indicated excess or wasted energy and any points beneath 1.0 indicate that not enough energy was genered.
</div>

```{r, message=FALSE}
tidy.data %>%
  ggplot(aes(x=day, y= generated / demand, color=location)) + 
  geom_point() +
  geom_smooth(se=FALSE) + 
  scale_color_brewer(palette='Set1') +
  theme(legend.position='bottom') +
  guides(color=guide_legend(nrow=3, override.aes=list(size=2))) +
  labs(x='Day in January 2018', 
       y='Proportion of Required Energy Generated',
       color='Location',
       title='California and New England Not Generating Enough Electricity to Meet Need'
       )
```

<div class='alert alert-info'>
California and New England are consistantly not producing enough energy to meet the needs of their populations. Every other measured locations is hovering near the required amount.
</div>











